# Market Matching and Causal Impact Inference
If you ever spent time in the field of marketing analytics, chances are that you have analyzed the causal impact of various events (interventions) using time series techniques. An event in this context could be a new TV or radio campaign, a major PR event, or some external event such as a new local competitor. These are all events where we cannot track the impact at the indidual customer level and hence have to analyze the impact from a bird's eye view – i.e., we have to analyze aggregated time series data, typically at the market level (e.g., DMA, state, etc.). Data science may be changing, but this is a use-case that has been around forever and is still very relavant.  

An intervention analysis usually involves two steps:

1. Find matching *control* markets for the *test* market where the event took place, using time series matching based on data prior to the event (pre period).
2. Analyze the causal impact of the event by comparing the observed data for the test and control markets following the event (post period), factoring in differences prior to the event. 

For step 1, the most straight-forward approach to this would be to use the Euclidian distance or correlation coefficient. However, this approach implicitly over-penalizes instances where the markets are shifted. Although it is preferable for test and control markets to be aligned consistently, occasional historical shifts should not eliminate control market candidates. For step 2, the traditional approach is a "diff in diff" model. However, this assumes  are i.i.d. and that the differences between the test and control markets is constant – assumptions that rarely hold true for time series data.

A better approach is to use *dynamic time warping* for the time series matching (step 1). This technique finds the distance along the *warping curve* – instead of the raw data – where the warping curve represents the best alignment between two time series within some user-defined constraint (time window). For the intervention analysis (step 2), a more robust approach is to use the method implemented in the `CausalImpact` package created by Kay Brodersen at Google (see [1]). The `CausalImpact` package constructs a synthetic baseline for the post-intervention period based on a structural time series model with linear regression components based on a set of matching control markets. 

## Note on the Workflow
We don't really need the dynamic time warping step to do execute this workflow; we can select markets directly while building the time series model in step 2. In fact, the `CausalImpact` package selects the most predictive markets for the structural time series model using spike-and-slab priors. This, we could jump directly to the `CausalImpact` package without pre-screening the markets. 

However, when dealing with a large set of candidate control markets it is often prudent to trim the list before selecting the final markets and we rarely need more than three markets to create robust synthetic control series. Moreover, basing the synthetic control on markets that have small distances to the test market boosts the face-validity of the analysis as these control markets tend to be similar in "size" and hence be recognized as viable controls.  

Ultimately, this is a matter of preference. The good news is that the `MarketMatching` package allows users to decide how many control markets should pass the filter in step 1.

# About MarketMatching Package

The `MarketMatching` package implements the workflow described above by providing a "wrapper" for the `dtw` and `CausalImpact` packages. R packages are a great way of implementing and documenting workflows.

Hence, the package does not provide any functionality that cannot be found in these packages, but rather simplifies the workflow by setting up the data and providing output that can easily be extracted for presentations. Summary of features:

* Minimal inputs required. The only strictly necessary inputs are the name of the test market (for inference) and the dates of the pre-period and post-period.
* Provides a data.frame with the best $K$ matches for all markets in the input dataset. $K$ can be defined by the user.
* Outputs all inference results as objects with intuitive names (e.g., "AbsoluteEffect" and "RelativeEffect").
* Calculates MAPE and Durbin-Watson for the pre-period. Shows how these statistics change when you alter the prior standard error of the local level term.
* Plots and outputs the actual data for the markets selected during the initial market matching.
* Plots and outputs actual versus predicted values.
* Plots the final local level term.
* Shows the average estimated coefficients for all the markets used in the linear regression component of the structural time series model.
* Allows the user to choose how many markets are sent to the slab-and-prior model.
* All plots are done in `ggplot2` and can easily be manipulated.

# How to Install
```{r, eval=FALSE}
library(devtools)
install_github("google/CausalImpact")
install_github("klarsen1/MarketMatching", "klarsen1")
```

# Example
```{r, echo = TRUE, message=FALSE, eval=FALSE}
library(MarketMatching)
##-----------------------------------------------------------------------
## Find the best matches (default is 5) for each airport time series
##-----------------------------------------------------------------------
library(MarketMatching)
data(weather, package="MarketMatching")
mm <- best_matches(data=weather, 
                   id_variable="Area", 
                   date_variable="Date", 
                   matching_variable="Mean_TemperatureF", 
                   parallel=TRUE,
                   start_match_period="2014-01-01",
                   end_match_period="2014-10-01")
##-----------------------------------------------------------------------
## Analyze causal impact of a made-up weather intervention in Copenhagen
## Since this is weather data this is a meaningless example and we should 
## expect no causal impact. This is just to demo the function.
##-----------------------------------------------------------------------
library(CausalImpact)
results <- MarketMatching::inference(matched_markets = mm, 
                                    test_market = "CPH", 
                                    end_post_period = "2015-10-01")
``` 

A view of the best matches data.frame generated by the best_matches() function:
```{r, echo = TRUE, message=FALSE, eval=FALSE}
knitr::kable(head(mm$BestMatches))
```

Plotting the absolute impact
```{r, echo = TRUE, message=FALSE, eval=FALSE}
results$PlotAbsoluteEffect
```

Plot actual observations for test market (CPH) versus the expectation
```{r, echo = TRUE, message=FALSE, eval=FALSE}
results$PlotActualVersusExpected
```

Store actual versus predicted in a data.frame
```{r, echo = TRUE, message=FALSE, eval=FALSE}
pred <- results$Predictions
knitr::kable(head(pred))
```

PLot the actual data for the test and control markets
```{r, echo = TRUE, message=FALSE, eval=FALSE}
results$PlotActuals
```

Check DW, MAPE and largest market coefficient for different values of the local level SE
```{r, echo = TRUE, message=FALSE, eval=FALSE}
results$PlotPriorLevelSdAnalysis
```

Store the coefficients in a data.frame
```{r, echo = TRUE, message=FALSE, eval=FALSE}
coeff <- results$Coefficients
knitr::kable(head(coeff))
```

# References
[1] CausalImpact version 1.0.3, Brodersen et al., Annals of Applied Statistics (2015). http://google.github.io/CausalImpact/

[2] Vignette for the `dtw` package: https://cran.r-project.org/web/packages/dtw/vignettes/dtw.pdf.