# Market Matching and Causal Impact Inference
If you ever spent time in the field of marketing analytics, chances are that you have analyzed the causal impact of various events (interventions) using time series techniques. An event could be a new TV or radio campaign, a major PR event, or some external event such as a new local competitor. These are all events where we cannot track the impact at the indidual customer level and hence have to analyze the impact from a bird's eye view – i.e., we have to analyze aggregated time series data, typically at the market level (e.g., DMA state, etc.).   

An intervention analysis usually involves two steps:

1. Find *control* market(s) for the *test* market(s) where the event took place, using time series matching based on data prior to the event (pre period).
2. Analyze the causal impact of the event by comparing the observed data for the test and control markets following the event (post period), factoring in differences prior to the event. 

The most straight-forward approach to this would be to use a simple Euclidian distance for step 1, and then a "diff in diff" model for step 2. However, this assumes that markets are always aligned historically and that the difference between the test and control markets is constant. Although we prefer markets to be aligned, shifts can occur and we do not want to overpenalize occasional deviations. Moreover, the gap between the test and control markets can fluctuate and this should be factored into the uncertainty of the post period analysis. 

A better approach is to use *dynamic time warping* for the time series matching and a Bayesian structural time series model to do the post period inference. The dynamic time warping can be achieved with the `dtw` package (available on CRAN) and the post period analysis can be achieved with the `CausalImpact` analysis created by Kay Brodersen at Google. In factm the MarketMatching package can be considered a "wrapper" for combining these two packages

In theory, we don't need a test and a control market to do an intervention analysis – we can use a structural time series model derived from the test group to establish the the baseline for the post period. However, introducing a control market as an input into the expectation for the post period promotes transparancy in terms of where the expectation comes from and does a nice job of capturing seasonality when history is limited or a short pre period is desired. 

# About MarketMatching Package

The `MarketMatching` package is designed to perform the two key steps needed to do an intervention analysis:

* For all markets in the input dataset, find the best control market using time series matching
* Given a test market and a matching control market (from above), analyze the causal impact of an intervention

# Simple Example
```{r, echo = TRUE, message=FALSE, eval=FALSE}
library(MarketMatching)
##-----------------------------------------------------------------------
## Find best matches for each airport time series
##-----------------------------------------------------------------------
library(MarketMatching)
data(weather, package="MarketMatching")
mm <- best_matches(data=weather, id="Area", 
                  date_variable="Date", 
                  matching_variable="Mean_TemperatureF", 
                  parallel=FALSE, 
                  start_match_period="2014-01-01",
                  end_match_period="2014-10-01")
head(mm$Distances)
subset(mm$Distances, Area=="CPH")
 
##-----------------------------------------------------------------------
## Analyze causal impact of a made-up weather intervention in Copenhagen
## Since this is weather data this is a meaningless example and we should 
## expect no causal impact. This is just to demo the function.
##-----------------------------------------------------------------------
results <- MarketMatching::inference(matched_markets = mm, test_market = "CPH", end_post_period = "2015-10-01")
 
## Plot the impact
results$PlotAbsoluteEffect
 
## Plot actual observations for test market (CPH) versus the expectation (based on the control)
results$PlotActualVersusFitted
```

# How to Install
```{r, eval=FALSE}
library(devtools)
install_github("google/CausalImpact")
install_github("klarsen1/MarketMatching", "klarsen1")
```

# References
CausalImpact version 1.0.3, Brodersen et al., Annals of Applied Statistics (2015). http://google.github.io/CausalImpact/

Vignette for the `dtw` package: https://cran.r-project.org/web/packages/dtw/vignettes/dtw.pdf.