---
title: "MarketMatching Package Vignette"
author: "Kim Larsen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MarketMatching Vignette}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
---

# Market Matching and Causal Impact Inference
If you ever spent time in the field of marketing analytics, chances are that you have analyzed the existence of a causal impact from a new TV campaign, a major PR event, or the entrance of a new local competitor using some sort of time series approach. The impact of such events cannot be tracked at the individual customer level and we do not have the luxury of clean test and control groups. Instead we have to analyze the impact from a bird's eye view using time series analysis at the market level (e.g., DMA, state, etc.). Data science may be changing at a fast pace but this is an old school use-case that is still very relevant.

An intervention analysis usually involves two steps:

1. Find matching *control* markets for the *test* market where the event took place using time series matching based on historical data prior to the event (the "pre period"").
2. Analyze the causal impact of the event by comparing the observed data for the test and control markets following the event (the "post period""), while factoring in differences between the markets prior to the event.

## A Traditional Approach
For the first step, the most straightforward approach is to use the Euclidian distance to find matching control markets. However, this approach implicitly over-penalizes instances where the markets are shifted. Although it is preferable for test and control markets to be aligned consistently, occasional historical shifts should not eliminate viable control market candidates.

For the second step, the traditional approach is a "diff in diff" analysis. This is typically a static regression model that evaluates the change in the difference between the test and control markets before and after the event. However, this assumes that observations are i.i.d. and that the differences between the test and control markets, before and after the event, are constant. Both assumptions rarely hold true for time series data.

## A More Flexible and Robust Approach
A better approach is to use *dynamic time warping* for the time series matching step. This technique finds the distance along the *warping curve* – instead of the raw data – where the warping curve represents the best alignment between two time series within some user-defined constraints. Note that the Euclidian distance is a special case of the warped distance.

For the intervention analysis the `CausalImpact` package created by Kay Brodersen at Google provides an approach that is more flexible and robust than the "diff in diff" model  (see [1]). The `CausalImpact` package constructs a synthetic baseline for the post-intervention period based on a Bayesian structural time series model that leverages *multiple* matching control markets as predictors as well as other features of the time series.

We can summarize this workflow as follows:

1. Find the best control markets for each market in the dataset using dynamic time warping. The user can define how many matches should be retained. 

2. Fit a Bayesian structural time series model that leverages the control markets identified in step 1. Based on this model, create a synthetic control series by producing a counterfactual prediction for the post period assuming that the event did not take place. Then calculate the difference between the synthetic control and the test market for the post-intervention period – which is the estimated impact of the event – and compare to the posterior interval to gauge uncertainty. 

### Notes on the Workflow
The dynamic time warping step is not a strictly necessary step to execute this workflow; we can select markets directly while building the time series model during step 2. In fact, the `CausalImpact` package selects the most predictive markets for the structural time series model using spike-and-slab priors. This, we could jump directly to the `CausalImpact` package without pre-screening the markets.

However, when dealing with a large set of candidate control markets it often turns out to be prudent to trim the list in advance, as opposed to relying solely on the variable selection process. In other words, pre-screening the markets is akin to wearing belt and suspenders; you're much less likely to get a strange and indefensible model. Moreover, basing the synthetic control on markets that have small *distances* to the test market boosts the face-validity of the analysis as these control markets will be similar in "size" and hence be recognized as viable controls by your business partners.  Ultimately, however, this is a matter of preference and the good news is that the `MarketMatching` package allows users to decide how many control markets should pass through step 1.

Last but not least, one can argue that the initial screening should be correlation-based and that distances are irrelevant – i.e., the "sizes" of the control markets do not matter as long as they are predictive. Again, this is a matter of preferences and `MarketMatching` allows the user to control whether the ranking is correlation-based or distance-based.

# About MarketMatching Package

The `MarketMatching` package implements the workflow described above by essentially providing an easy-to-use "wrapper" for the `dtw` and `CausalImpact`. The function `best_matches()` finds the best control markets for each market by looping through all viable candidates in a parallel fashion and then ranking by distance and/or correlation. The resulting output object can then be passed to the `inference()` function which then analyzes the causal impact of an event using the control markets identified by `best_matches()`. 

Hence, the package does *not* provide any core functionality that cannot be found in these packages but it simplifies the workflow of using `dtw` and `CausalImpact` together *and* provides charts and data that are easy to manipulate. R packages are simply a great way of implementing and documenting workflows.

## Summary of features:

* Minimal inputs required. The only strictly necessary inputs are the name of the test market (for inference), the dates of the pre-period and post-period and, of course, the data.
* Provides a data.frame with the best matches for all markets in the input dataset. The number of matches can be defined by the user.
* Outputs all inference results as objects with intuitive names (e.g., "AbsoluteEffect" and "RelativeEffect").
* Checks the quality of the input data and eliminates "bad" markets.
* Calculates MAPE and Durbin-Watson for the pre-period. Shows how these statistics change when you alter the prior standard error of the local level term.
* Plots and outputs the actual data for the markets selected during the initial market matching.
* Plots and outputs actual versus predicted values.
* Plots the final local level term.
* Shows the average estimated coefficients for all the markets used in the linear regression component of the structural time series model.
* Allows the user to choose how many markets are sent to the slab-and-prior model.
* All plots are done in `ggplot2` and can easily be extracted and manipulated.

# How Dynamic Time Warping Works
Let's say we have two time series denoted by $X=(x_1, \ldots, x_n)$ and $Z=(z_1, \ldots, z_m)$, where $X$ is the test market (also called the *reference index*) and $Z$ is the control market (also called the *query index*). Note that, although $m$ and $n$ do not need to be equal, `MarketMatching` forces $m=n$. We'll denote the common length by $T$.

In order to calculate the distance between these two time series, the first step is to create the *warping curve* $\phi(t) = (\phi_x(t), \phi_z(t))$. The goal of the warping curve is to remap the *indexes* of the original time series – through the *warping functions* $\phi_x(t)$ and $\phi_z(t)$ – such that the remapped series are as similar as possible, where similarity is defined by

$$ D(X,Z) = \frac{1}{M_{\phi}} \sum_{i=1}^T d(\phi_x(t), \phi_z(t))m_{\phi}(t). $$

Here $d(\phi_x(t), \phi_z(t))$ is the local distance between the remapped data points and at index $t$, $m_{\phi}(t)$ is the per-step weight and $M_{\phi}$ is an optional normalization constant (only relevant if $m \neq n$). The per-step weights are defined by the "step pattern" which provides a flexible approach to control the slopes along the warping curve. For more details on this, and everything else related to dynamic time warping in R, see [2].

Thus, the goal is essentially to find the warping curve, $\phi$ such that $D(X,Z)$ is minimized. Standard constraints for this optimization problem include:

* Monotonicity: ensures that the ordering of the indexes of the time series are preserved – i.e., $\phi_x(t+1) > \phi_x(t)$.
* Warping limits: limits the length of the permissible steps. The `MarketMatching` package specifies the well known the Sakoe-Chiba band (when calling `dtw`) which allows the user to specify a maximum allowed time difference between two matched data points. This can be expressed as  $|\phi_x(t)-\phi_z(t)<L$, where $L$ is the maximum allowed difference. For example, if your data is weekly and $L=2$, we would never allow the warping curve to skip more than four weeks.

## Example
To see how this works, consider the following example. We'll use the weather dataset included with the `MarketMatching` package and use the first 10 days from the Copenhagen time series as the test market and San Francisco as the control market (query series).

Note that the code in this example is *not* needed to run the `MarketMatching` package, the package will set it up for you. This is simply to demonstrate the details behind the scene.

First, let's look at the warping limits imposed by the Sakoe-Chiba band with $L=2$:
```{r, echo = TRUE, message=FALSE, eval=TRUE}
library(MarketMatching)
library(dtw)
data(weather, package="MarketMatching")

cph <- subset(weather, Area=="CPH")$Mean_TemperatureF[1:10]
sfo <- subset(weather, Area=="SFO")$Mean_TemperatureF[1:10]
cph
sfo

align <- dtw(cph, sfo, window.type=sakoeChibaWindow, window.size=2, keep=TRUE)
```

```{r, echo = TRUE, message=FALSE, eval=TRUE, fig.width=7, fig.height=5}
dtwWindow.plot(sakoeChibaWindow, window.size=2, reference=10, query=10)
```

This shows that, as expected, the band is a symmetric constraint around the diagonal.

Next, let's look at the alignment between the two time series. The following code shows the two time series as well as how data points are connected:

```{r, echo = TRUE, message=FALSE, eval=TRUE, fig.width=7, fig.height=5}
plot(align,type="two", off=1)
```

This shows that the two cities are not well aligned naturally (not surprising), and that some reference values are mapped to four different query values (the most allowed).

It also helps to look at the actual cost matrix and the optimal alignment path that leads to the minimal distance.

```{r, echo = TRUE, message=FALSE, eval=TRUE, fig.width=7, fig.height=5}
lcm <- align$localCostMatrix
image(x=1:nrow(lcm),y=1:ncol(lcm),lcm,xlab="CPH Index",ylab="SFO Index")
text(row(lcm),col(lcm),label=lcm)
lines(align$index1,align$index2)
```

The cells represent the local distances between the two time series. The total cost (distance) can be computed by multiplying the distances by their respective weights and then summing up along the optimal path. This can be illustrated by the cumulative cost matrix:

```{r, echo = TRUE, message=FALSE, eval=TRUE, fig.width=7, fig.height=5}
lcm <- align$costMatrix
image(x=1:nrow(lcm),y=1:ncol(lcm),lcm,xlab="CPH Index",ylab="SFO Index")
text(row(lcm),col(lcm),label=lcm)
lines(align$index1,align$index2)
```
Note that with the default step pattern diagonal steps are weighted by a factor of 2 while the other steps receive a weight of 1. This means that a two-step "detour" will cost as much as a direct, diagonal step – assuming all local distances are equal. Thus, this step pattern allows for as much stretching of the warping curve as possible (as long as it is within the specified band). In this specific case, this means that we can match a single element of the query index (CPH) to as much as four points of the reference index (ZRH). If we want to enforce a more diagonal path, we can specify `StepPattern=symmetric1` in the `dtw` call. For a deeper explanation of how step patterns work, see [2].

The distance equals 195, which is a fairly large number given that the distance between Zurich and Copenhagen is only 44:

```{r, echo = TRUE, message=FALSE, eval=TRUE, fig.width=7, fig.height=5}
zrh <- subset(weather, Area=="ZRH")$Mean_TemperatureF[1:10]
match <- dtw(cph, zrh, window.type=sakoeChibaWindow, window.size=2, keep=TRUE)
lcm <- match$costMatrix
match$distance
image(x=1:nrow(lcm),y=1:ncol(lcm),lcm,xlab="CPH Index",ylab="ZRH Index")
text(row(lcm),col(lcm),label=lcm)
lines(match$index1,match$index2)
```

# How Does Intervention Inference Work?
As mentioned, the `MarketMatching` package utilizes the `CausalImpact` package written by Kay Brodersen at Google (see [1]) to do the post period inference. 
Here is how it works at a high level:

1. Fit a Bayesian structural time series model using data prior to the pre-intervention period. The model can include the control markets as linear regression components with spike-and-slab priors.

2. Based on this model, generate counterfactual predictions for the post-intervention period assuming that the intervention did not take place.

3. In a pure Bayesian fashion, leverage the counterfactual predictions to quantify the causal impact of the intervention.

This approach has a number of benefits over the "diff in diff" approach: First, using a structural time series model allows us to capture for latent evolutions of the test market that cannot be explained by known trends or events. Second, estimating control markets effects with Bayesian priors captures the uncertainty of the relationship between the test market and the control markets. This is critical as it ensures that the counterfactual predictions are not rigidly relying on historical relationships between the test and control markets that may be carry large standard errors. Moreover the spike-and-slab priors help avoid over fitting by promoting a sparseness during market (variable) selection.

As a result, this approach produces robust counterfactual expectations for the post period that factors in uncertainties in historical market relationships as well as unobserved trends. Moreover, we can calculate posterior intervals through sampling to gauge confidence in the magnitude of causal impact and estimate the posterior probability that the causal impact is non-existent. The "diff in diff" approach does not provide this level of flexibility and does not handle parameter-uncertainty nearly as well.

## Some Technical Details
When `MarketMatching` calls `CausalImpact` the following structural time series model (state space model) is created for the pre-intervention period:

$$ Y_t = \mu_t + x_t \beta + e_t, e_t \sim N(0, \sigma^2_e) $$
$$ \mu_{t+1} = \mu_t + \nu_t, \nu_t \sim N(0, \sigma^2_{\nu}) $$

Here $x_t$ denote the control markets and $\mu_t$ is the *local level* term. The local level term defines how the latent state evolves over time and hence is often referred to as the *unobserved trend*. The linear regression term, $x_t\beta$, averages over the selected control markets. See [1] and [3] for more details.

Once this model is in place, we can create a synthetic control series by predicting the values for the post period, and compare to the actual values. In order to gauge believability of the difference between the control series and the observed values, posterior intervals can be created through sampling in a pure Bayesian fashion. We can also compute the tail probability of a non-zero impact. The `CausalImpact` package conveniently calculates the posterior inference for us.

### About Spike-and-Slab Priors
The `CausalImpact` package applies spike-and-slab priors to the coefficients of the linear regression terms. In short, the spike-and-slab prior consist of two parts:

The spike part governs a market's probability of being chosen for the model (i.e., having a non-zero coefficient). This is typically a product of independent Bernoulli distributions (one for each variable), where the parameters (probability of getting chosen) can be set according to the expected model size. The slab part is a wide-variance Gaussian prior that shrinks the non-zero coefficients toward some value (usually zero).

This approach is a powerful way of reducing a large set of correlated markets into a parsimonious model that averages of a smaller set of markets. Moreover, since the market coefficients follow random distributions, we can incorporate the uncertainties of the historical relationships into the model as opposed to relying on a rigid encoding based on fixed coefficients when forecasting.

For more details in spike-and-slab priors, see [1].

### How to Select the Local Level Standard Error
There's no perfectly scientific way of choosing the standard error (SE) of the local level term. Using a pure goodness-of-fit-based measures is not meaningful as a higher SE will always guarantee a better historical fit and the goal of the model is not to fit the data in the post-intervention period (i.e., we cannot use the post intervention period as a hold-out sample).

Here are some tips to deciding the standard error:

* Note that larger values of the standard error leads to wider posterior forecast intervals and hence inconclusive results. Thus, choosing a large value "to be safe" is not always a prudent choice.

* If you know a priori that the series is volatile due to unexplained noise, choose 0.1.

* Try different values of the standard error, and check the MAPE and Durbin-Watson statistic. The MAPE measures the historical fit and the Durbin-Watson statistic measures the level of autocorrelation in the residuals. This shows you the tradeoff between a larger standard error versus fit and ill-behaved residuals. Note that the Durbin-Watson statistic should be as close to 2 as possible. We want to choose a standard error that is as small as possible in order rely more on the predictive value coming from the control markets, but not at any cost. The `MarketMatching` package produces charts that help make this tradeoff.

* When you cannot make a decision, choose 0.01.

### Note on Structural Time Series Models
The `CausalImpact` package can fit much more complicated models with seasonal terms and dynamic coefficients for the linear regression component. The class of models deployed by the `CausalImpact` package is flexible and, in fact, all ARIMA models can be converted into a structural model. For example, the ARIMA(0,1,1) model

$$ (1 − B)y_t = (1 − \rho B)a_t, $$

where $B$ is the backshift operator, $\rho$ is the AR(1) coefficient, and $a_t$ are the residuals, can be recast as the structural model with a local level term (the model described above). However, the structural model is more transparent by no operating in a differenced space and its Bayesian nature allows us to control the variance of the local level term and hence reduce the variance of the forecasted posterior values.

# How to Install
```{r, eval=FALSE}
library(devtools)
install_github("google/CausalImpact")
install_github("klarsen1/MarketMatching")
```

# Example
The dataset supplied with the package has daily temperature readings for 20 areas (airports) for 2014. The dataset is a stacked time series (panel data) where reach row represent a unique combination of date and area. It has three columns: Area, date, and the average temperature reading for the day in Fahrenheit.

This is *not* a great dataset to demonstrate intervention inference as humans cannot affect the weather in the short term (long term is a different story). We'll merely use the data to demonstrate the features.

```{r, echo = TRUE, message=FALSE, eval=TRUE}
##-----------------------------------------------------------------------
## Find the best matches (default is 5) for each airport time series
##-----------------------------------------------------------------------
library(MarketMatching)
data(weather, package="MarketMatching")
mm <- best_matches(data=weather,
                   id_variable="Area",
                   date_variable="Date",
                   matching_variable="Mean_TemperatureF",
                   parallel=TRUE,
                   start_match_period="2014-01-01",
                   end_match_period="2014-10-01")
##-----------------------------------------------------------------------
## Analyze causal impact of a made-up weather intervention in Copenhagen
## Since this is weather data this is a meaningless example and we should
## expect no causal impact. This is just to demo the function.
##-----------------------------------------------------------------------
library(CausalImpact)
results <- MarketMatching::inference(matched_markets = mm,
                                    test_market = "CPH",
                                    end_post_period = "2015-10-01")
```

A view of the best matches data.frame generated by the best_matches() function:
```{r, echo = TRUE, message=FALSE, eval=TRUE, results='asis'}
knitr::kable(head(mm$BestMatches))
```

Plot actual observations for test market (CPH) versus the expectation. It looks like CPH deviated from its expectation during the winter:
```{r, echo = TRUE, message=FALSE, eval=TRUE, fig.width=7, fig.height=5}
results$PlotActualVersusExpected
```

Plot the cumulative impact. The posterior interval includes zero as expected, which means that the cumulative deviation is likely noise:
```{r, echo = TRUE, message=FALSE, eval=TRUE, fig.width=7, fig.height=5}
results$PlotCumulativeEffect
```

Although it looks like some of the dips in the *pointwise* effects toward the end of the post period seem to be truly negative:
```{r, echo = TRUE, message=FALSE, eval=TRUE, fig.width=7, fig.height=5}
results$PlotPointEffect
```

Store the actual versus predicted values in a data.frame:
```{r, echo = TRUE, message=FALSE, eval=TRUE, results='asis'}
pred <- results$Predictions
knitr::kable(head(pred))
```

PLot the actual data for the test and control markets:
```{r, echo = TRUE, message=FALSE, eval=TRUE, fig.width=7, fig.height=5}
results$PlotActuals
```

Check DW, MAPE and largest market coefficient for different values of the local level SE. It looks like it will be hard to get a DW statistic close to 2, although our model may benefit from a higher local level standard error than the default of 0.01:
```{r, echo = TRUE, message=FALSE, eval=TRUE, fig.width=7, fig.height=5}
results$PlotPriorLevelSdAnalysis
```

Store the coefficients in a data.frame. STR (Stuttgart) receives the highest weight when predicting the weather in Copenhagen:
```{r, echo = TRUE, message=FALSE, eval=TRUE, results='asis'}
coeff <- results$Coefficients
knitr::kable(head(coeff))
```

# References
[1] CausalImpact version 1.0.3, Brodersen et al., Annals of Applied Statistics (2015). http://google.github.io/CausalImpact/

[2] Vignette for the `dtw` package: https://cran.r-project.org/web/packages/dtw/vignettes/dtw.pdf.

[3] Predicting the Present with Bayesian Structural Time Series, Steven L. Scott and Hal Varian, http://people.ischool.berkeley.edu/~hal/Papers/2013/pred-present-with-bsts.pdf.

